---
title: "PML_CourseProject"
author: "Nazmi Anik"
date: "June 5, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction  

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).  

## Loading the Data  

The data for this project come from [this source](http://groupware.les.inf.puc-rio.br/har).  
```{r load, echo=TRUE}
url_train <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
url_test <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

if(!file.exists('.\\data\\training.csv')){
    download.file(url=url_train, ".\\data\\training.csv")
    download.file(url=url_test, ".\\data\\testing.csv")
}

trainingRaw <- read.csv(".\\data\\training.csv", header=TRUE, sep = ",", na.strings = c("", "NA", "#DIV/0!"))
testingRaw <- read.csv(".\\data\\testing.csv", header=TRUE, sep = ",", na.strings = c("", "NA", "#DIV/0!"))
```

## Cleaning the Data  

Now, let's clean the dataset a little bit. Let's remove all the columns where the only values are "NA".  
```{r NAremove, echo=TRUE}
testing <- testingRaw[ , ! apply( testingRaw , 2 , function(x) all(is.na(x)) ) ]
dim(testingRaw)-dim(testing)
NAindex <- apply(testingRaw,2,function(x) {sum(is.na(x))}) 
training <- trainingRaw[,which(NAindex == 0)]
```
Based on the [class discussion](https://www.coursera.org/learn/practical-machine-learning/discussions/weeks/4/threads/1gqWVBxTEea3JA6-1_N9Kw/replies/CPHPZCYhEeaDFg6CjeDgQw), it is assumed that there is enough data that is non-missing in both datasets. Therefore using the method in [source](http://stackoverflow.com/a/15968711), we removed the columns that are "NA" in test data, also in the training data.  
Let's also take out the index column, the first column in both data sets.
```{r subset, echo=TRUE}
training <- training[,2:60]
testing <- testing[,2:60]
```
## Training 

### Getting the Training Set Ready
Let's separate the training data into training and test sets to work on our model fit.  
```{r model, echo=TRUE, message=FALSE}
library(caret)
library(rattle)
set.seed(1234)
inTrain = createDataPartition(training$classe, p=0.7, list=FALSE)
train1 = training[ inTrain,]
test1 = training[-inTrain,]
```

### Training with Decision Tree
We will first try the decision trees to fit our data.
```{r tree, echo=TRUE, message=FALSE, fig.height=4, fig.width=6}
modTree <- train(classe ~., method="rpart", data=train1)
fancyRpartPlot(modTree$finalModel)
```

